How to obtain the redshift distribution from probabilistic redshift estimates

### 2017-06-28 research/paper/draft-editing.tex
merged test cases
fixed typos

% It would be easy to estimate $\vec{\phi}$ given $\{z_{i}\}_{N}$, however, what we actually have is photometric data $\{\vec{d}_{i}\}_{N}$ which we used to derive photo-$z$ PDFs.  Because a photo-$z$ PDF is a probability distribution conditioned on the photometric data, it must be interpreted as a posterior probability distribution.

% In addition to the data, the posterior probability distribution is conditioned on other information about how it was made, which is in general specific to the method that produced it.  This information only influences the posterior by its implicit distribution over redshifts, which we call the interim prior over redshift.  The interim prior, like the redshift distribution we wish to infer, has some functional form with some parameters, but since the redshift distribution and interim prior need not share the same parametrization, we will use $\vec{\alpha}$ to represent the parameters for the interim prior.  Thus, the posterior probability that is often oversimplified as "$p(z_{i})$" is really the interim posterior probability $p(z_{i} | \vec{d}_{i}, \vec{\alpha})$.  Note that we will assume that all galaxies whose photo-$z$ interim posteriors are derived by a single method share the same interim prior $p(z | \vec{\alpha})$; depending on the method, this may not always be true.

@@

According to Bayes' rule, the posterior distribution of interest is
\begin{align}
\log p(\vec{\phi} | \{\vec{d}_{i}\}) &\propto \log p(\{\vec{d}_{i}\} | \vec{\phi}) + \log p(\vec{\phi}),
\end{align}
where $p(\vec{\phi})$ is the \textit{hyperprior} probability distribution over possible values of the hyperparameters.  The hyperprior is a choice made by those performing the inference and is not to be confused with the interim prior parameters that influence the photo-$z$ PDF data product.

Focusing on the likelihood term $p(\{\vec{d}_{i}\} | \vec{\phi})$, we employ the power of the hierarchical model through \textit{marginalization} of the unobserved redshift parameters by integrating over them to obtain 
\begin{align}
\log[p(\{\vec{d}_{i}\} | \vec{\phi})] &= \log[\int\exp[\log[p(\{\vec{d}_{i}\} | \{z_{i}\})]\\
& + \log[p(\{z_{i}\} | \vec{\phi})]]\ d\{z_{i}\}].
\end{align}

Given a catalog of photo-$z$ interim posteriors $\{p(z_{i} | \vec{d}_{i}, \vec{\alpha})\}$

To make use of the photo-$z$ interim priors, it is absolutely essential that $\vec{\alpha}$ be known.

This framework entails a number of choices and assumptions that must be addressed explicitly.

\begin{enumerate}
	\item The chosen functional form with parameters $\vec{\phi}$ must actually be able to describe the true redshift distribution.
	\item The photometric data and redshift for each galaxy is independent of the photometric data and redshifts of all other galaxies.
	\item There is one interim prior with parameters $\vec{\phi}^{*}$ shared among all galaxies in the survey.
	\item We must choose a hyperprior probability distribution $p(\vec{\phi})$ over the hyperparameters.
	\item While we advocate for the approach of hierarchical inference, the probabilistic graphical model presented here is not the only one that could be proposed.
\end{enumerate}

@@

It is useful to translate some popular existing methods for deriving $n(z)$ from photo-$z$ PDFs into the mathematical framework of Sec. \ref{sec:model}.

@@

%\appendix{}

>>>>

@@

According to Bayes' rule, the posterior distribution of interest is
\begin{align}
\label{eq:hyper_bayes}
\log[p(\vec{\phi} | \{\vec{d}_{i}\})] &\propto \log[p(\{\vec{d}_{i}\} | \vec{\phi})] + \log[p(\vec{\phi})] ,
\end{align}
where $p(\vec{\phi})$ is the \textit{hyperprior} probability distribution over possible values of the hyperparameters.  The hyperprior is a choice made by those performing the inference and is not to be confused with the interim prior parameters that influence the photo-$z$ PDF data product.

Focusing on the likelihood term $p(\{\vec{d}_{i}\} | \vec{\phi})$, we employ the power of the hierarchical model through \textit{marginalization} of the unobserved redshift parameters by integrating over them to obtain 
\begin{align}
\label{eq:marginalize}
\begin{split}
\log[p(\{\vec{d}_{i}\} | \vec{\phi})] = & \log\left[\int\exp\left[\log[p(\{\vec{d}_{i}\} | \{z_{i}\})]\right.\right. \\
& \left.\left.+ \log[p(\{z_{i}\} | \vec{\phi})]\right]\ d\{z_{i}\}\right] .
\end{split}
\end{align}
The redshifts $\{z_{i}\}$ are statistically independent of each other, and each galaxy's photometry depends solely on its own redshift, so we may invoke statistical independence to get
\begin{align}
\label{eq:independence_hyper}
\log[p(\{z_{i}\} | \vec{\phi})] &= \sum_{i}^{N} \log[p(z_{i} | \vec{\phi})]
\end{align}
and
\begin{align}
\label{eq:independence}
\log[p(\{\vec{d}_{i}\} | \{z_{i}\})] &= \sum_{i}^{N} \log[p(\vec{d}_{i} | z_{i})] .
\end{align}
Having chosen the functional form mapping $\vec{\phi}$ to $n(z)$, the terms on the righthand side of Eq. \ref{eq:independence_hyper} are known.  However, the terms on the righthand side of Eq. \ref{eq:independence} are likelihoods, and the photo-$z$ PDFs we have are interim posteriors.

To transform the interim posteriors we have into the likelihoods we need, we start with the vacuously true statement
\begin{align}
\begin{split}
\label{eq:unity}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})]\\
& - \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})].
\end{split}
\end{align}
We expand the last term using Bayes' Rule to get
\begin{align}
\begin{split}
\label{eq:bayes}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})]\\
& - \log[p(\vec{d}_{i} | z_{i}, \vec{\phi}^{*})]\\
& + \log[p(\vec{d}_{i} | \vec{\phi}^{*})] - \log[p(z_{i} | \vec{\phi}^{*})].
\end{split}
\end{align}
Fig. \ref{fig:pgm} has no direct connection between $\vec{d}_{i}$ and $\vec{\phi}$, so we can break up $\log[p(\vec{d}_{i} | z_{i}, \vec{\phi}^{*})]$ as follows:
\begin{align}
\begin{split}
\label{eq:split}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})]\\
& - \log[p(\vec{d}_{i} | z_{i})] - \log[p(\vec{d}_{i} | \vec{\phi}^{*})]\\
& + \log[p(\vec{d}_{i} | \vec{\phi}^{*})] - \log[p(z_{i} | \vec{\phi}^{*})].
\end{split}
\end{align}
Canceling terms, we find
\begin{align}
\label{eq:cancel}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})] - \log[p(z_{i} | \vec{\phi}^{*})].
\end{align}

Now we combine Eqs. \ref{eq:hyper_bayes}, \ref{eq:marginalize}, \ref{eq:independence_hyper}, \ref{eq:independence}, and \ref{eq:cancel} to arrive at 
\begin{align}
\begin{split}
\label{eq:final}
\log[p(\vec{\phi} | \{\vec{d}_{i}\})] \propto& \log[p(\vec{\phi})] + \log\left[\int \exp\left[\sum_{i}^{N} \left(\right.\right.\right.\\
& \left.\left.\left. + \log[p(z_{i} | \vec{\phi})] - \log[p(z_{i} | \vec{\phi}^{*})]\right.\right.\right.\\
& \left.\left.\left. + \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})]
\right)\right] d\{z_{i}\}\right].
\end{split}
\end{align}
Thus, if we have a catalog of photo-$z$ interim posteriors $\{p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})\}$, the interim prior parameters $\vec{\phi}^{*}$, and a prior distribution $p(\vec{\phi})$, we may find the posterior $\log[p(\vec{\phi} | \{\vec{d}_{i}\})]$ on the redshift distribution $n(z)$.

This framework entails a number of choices and assumptions that must be addressed explicitly:
\begin{enumerate}
	\item The chosen functional form of $n(z)$ with parameters $\vec{\phi}$ must be capable of describing the true redshift distribution.
	\item The photometric data and redshift for each galaxy must be independent of the photometric data and redshifts of all other galaxies.
	\item There is one interim prior shared among all galaxies in the survey, and its parameters with known parameters $\vec{\phi}^{*}$ are known.
	\item We must choose a hyperprior probability distribution $p(\vec{\phi})$ over the hyperparameters.
	\item While we advocate for the approach of hierarchical inference, the probabilistic graphical model presented here is not the only one that could be proposed.
\end{enumerate}

@@

It is useful to translate some popular existing methods for deriving $n(z)$ from photo-$z$ PDFs into the mathematical framework of Sec. \ref{sec:model}.  We briefly discuss the conditions under which they are equivalent to Eq. \ref{eq:final}, which are elaborated upon in the Appendix.

\subsubsection{Stacking}
\label{sec:stacking}

The most common way to combine photo-$z$ interim posteriors into an estimator $\hat{n}_{stack}(z)$ for $n(z)$ is to "stack" them, which corresponds to
\begin{align}
\label{eq:stack}
\hat{n}^{stack}(z) &= \frac{1}{N}\ \sum_{i}^{N} \exp\left[\log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})]\right].
\end{align}

\subsubsection{Point Estimation}
\label{sec:point_estimation}

In some cases, estimators $\hat{n}^{point}(z)$ are obtained by assuming the photo-$z$ PDFs are effectively delta functions by reducing them to point estimates according to
\begin{align}
\label{eq:delta}
p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*}) &\approx \delta(z, \hat{z}_{i}^{point})
\end{align}
before applying Eq. \ref{eq:stack}.  The most popular redshift point estimators are the mean $\hat{z}^{mean}_{i}$, median $\hat{z}^{median}_{i}$, and mode $\hat{z}^{mode}_{i}$ of the original photo-$z$ interim posterior.

@@

\appendix{}
%showing stacking is always too broad
